Windowed least-squares model reduction for dynamical systems

### 2019-10-23 cavity.tex
updating cav figs

\subsection{Cavity flow}
The next numerical example considers collocated ROMs\footnote{Collocation is a form of hyper-reduction which requires sampling the full-order model at only select grid points.} of a viscous, compressible flow in a two-dimensional cavity. 
The flow is described by the two-dimensional compressible Navier-Stokes equations,
\begin{equation}\label{eq:compressible_ns}
\frac{\partial \mathbf{u}}{\partial t} + \nabla \cdot \big( \mathbf{F}(\mathbf{u} ) - \mathbf{F}_v (\mathbf{u},\nabla \mathbf{u} )      \big) =\bz,
\end{equation}
where $\mathbf{u}: [0,T] \times \Omega \rightarrow \RR{4}$ comprise the density, $\mathsf{x}_1$ and $\mathsf{x}_2$ momentum, and total energy. The terms $\mathbf{F}$ and $ \mathbf{F}_v$ are the inviscid and viscous fluxes, respectively. For a two-dimensional flow, the state vector and inviscid fluxes are
$$
\mathbf{u} = \begin{Bmatrix}
\rho \\ \rho u_1 \\ \rho u_2 \\ \rho E \end{Bmatrix}, \qquad \mathbf{F}_{1} = \begin{Bmatrix} \rho u_1 \\ \rho u_1^2 +      p \\ \rho u_1 u_2 \\ u_1(E + p) \end{Bmatrix}, 
\qquad \mathbf{F}_{2} = \begin{Bmatrix} \rho u_2 \\ \rho u_1 u_2  \\ \rho u_2^2 + p \\ u_2(E + p) \end{Bmatrix}.
$$
The viscous fluxes are given by
$$
\qquad \mathbf{F}_{v_1} = \begin{Bmatrix} 0 \\ \tau_{11} \\ \tau_{12}  \\ u_j \tau_{j1} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_1}  \end{Bmatrix}, 
\qquad \mathbf{F}_{v_2} = \begin{Bmatrix} 0 \\ \tau_{21} \\ \tau_{22}  \\ u_j \tau_{j2} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_2}  \end{Bmatrix},
$$
where $\mu \in \RRplus$ is the viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
We assume a Newtonian fluid, which leads to a viscous stress tensor of the form
\begin{equation*}
\tau_{ij} = 2\mu S_{ij},
\end{equation*}
where
\begin{equation*}
 S_{ij} = \frac{1}{2} \big( \frac{\partial u_i}{\partial \mathsf{x}_j} + \frac{\partial u_j}{\partial \mathsf{x}_i} \big) - \frac{1}{3} \frac{\partial      u_k}{\partial \mathsf{x}_i} \delta_{ij}.
\end{equation*}
We close the Navier--Stokes equations with a constitutive relationship for a calorically perfect gas,
$$p = (\gamma - 1)( \rho E - \frac{1}{2} \rho u_1^2 - \frac{1}{2} \rho u_2^2 \big),$$
where $\gamma = 1.4$ is the heat-capacity ratio.

Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions. We note that the Reynolds number is defined as $\text{Re} = \rho \norm{\mathbf{v}_{\infty}} L / \mu$ with a characteristic length set at $L=1$ and $\mathbf{v} = [u_1 \; \; u_2]^T$, the speed of sound by $a_{\infty} = \sqrt{\gamma p_{\infty}/\rho}$, and $\infty$ subscripts refer to free-stream conditions. We employ free-stream boundary conditions at the inlet, outlet, and top wall of the cavity. We enforce no-slip boundary conditions 
on the bottom wall of the cavity.

@@

Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROM\ with $\Delta T = 0.5$ for the time instance $t = 5.0$. LSPG is observed to exhibit artificial oscillations; the Gauss-Newton method fails to converge at $t \approx 16.0$. The \methodAcronymROM\ at $\Delta T = 0.5$ is able to capture the important features of the flow, including the points of flow separation at the start and end of the ramp, and remains stable for the entire time interval.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_fom_t5_basis2.pdf}
\caption{FOM} 
\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c2_t10.pdf}
%\caption{\methodAcronym, $\Delta T = 0.2$} 
%\end{subfigure}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_lspg_t5_basis2.pdf}
\caption{LSPG} 
\end{subfigure}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c5_t5_basis2.pdf}
\caption{\methodAcronym, $\Delta T = 0.5$} 
\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c10_t10.pdf}
%\caption{\methodAcronym, $\Delta T = 2.0$} 
%\end{subfigure}
\caption{Field solutions for the vorticity at $t = 5.0$}
\label{fig:cav_snapshots}
\end{center}
\end{figure}

>>>>

\subsection{Cavity flow}
The next numerical example considers collocated ROMs\footnote{Collocation is a form of hyper-reduction which requires sampling the full-order model at only select grid points.} of a viscous, compressible flow in a two-dimensional cavity. 
The flow is described by the two-dimensional compressible Navier-Stokes equations,
\begin{equation}\label{eq:compressible_ns}
\frac{\partial \mathbf{u}}{\partial t} + \nabla \cdot \big( \mathbf{F}(\mathbf{u} ) - \mathbf{F}_v (\mathbf{u},\nabla \mathbf{u} )      \big) =\bz,
\end{equation}
where $\mathbf{u}: [0,T] \times \Omega \rightarrow \RR{4}$ comprise the density, $\mathsf{x}_1$ and $\mathsf{x}_2$ momentum, and total energy. The terms $\mathbf{F}$ and $ \mathbf{F}_v$ are the inviscid and viscous fluxes, respectively. For a two-dimensional flow, the state vector and inviscid fluxes are
$$
\mathbf{u} = \begin{Bmatrix}
\rho \\ \rho u_1 \\ \rho u_2 \\ \rho E \end{Bmatrix}, \qquad \mathbf{F}_{1} = \begin{Bmatrix} \rho u_1 \\ \rho u_1^2 +      p \\ \rho u_1 u_2 \\ u_1(E + p) \end{Bmatrix}, 
\qquad \mathbf{F}_{2} = \begin{Bmatrix} \rho u_2 \\ \rho u_1 u_2  \\ \rho u_2^2 + p \\ u_2(E + p) \end{Bmatrix}.
$$
The viscous fluxes are given by
$$
\qquad \mathbf{F}_{v_1} = \begin{Bmatrix} 0 \\ \tau_{11} \\ \tau_{12}  \\ u_j \tau_{j1} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_1}  \end{Bmatrix}, 
\qquad \mathbf{F}_{v_2} = \begin{Bmatrix} 0 \\ \tau_{21} \\ \tau_{22}  \\ u_j \tau_{j2} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_2}  \end{Bmatrix},
$$
where $\mu \in \RRplus$ is the dynamic viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
We assume a Newtonian fluid, which leads to a viscous stress tensor of the form
\begin{equation*}
\tau_{ij} = 2\mu S_{ij},
\end{equation*}
where
\begin{equation*}
 S_{ij} = \frac{1}{2} \big( \frac{\partial u_i}{\partial \mathsf{x}_j} + \frac{\partial u_j}{\partial \mathsf{x}_i} \big) - \frac{1}{3} \frac{\partial      u_k}{\partial \mathsf{x}_i} \delta_{ij}.
\end{equation*}
We close the Navier--Stokes equations with a constitutive relationship for a calorically perfect gas,
$$p = (\gamma - 1)( \rho E - \frac{1}{2} \rho u_1^2 - \frac{1}{2} \rho u_2^2 \big),$$
where $\gamma = 1.4$ is the heat-capacity ratio.

Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions. The Reynolds number is defined as $\text{Re} = \rho_{\infty} \norm{\mathbf{v}_{\infty}} L / \mu$ with a characteristic length set at $L=1$ and $\mathbf{v} = [u_1 \; \; u_2]^T$, the speed of sound is defined by $a_{\infty} = \sqrt{\gamma p_{\infty}/\rho_{\infty}}$, and $\infty$ subscripts refer to free-stream conditions. We employ free-stream boundary conditions at the inlet, outlet, and top wall of the cavity. We enforce no-slip boundary conditions 
on the bottom wall of the cavity.

@@

Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROMs\ with $\Delta T = 0.5,1.0$ for the time instance $t = 5.0$. LSPG is observed to exhibit artificial oscillations; the Gauss-Newton method fails to converge at $t \approx 16.0$. The \methodAcronymROM\ at $\Delta T = 0.5$ is able to capture the important features of the flow, including the points of flow separation at the start and end of the ramp, and remains stable for the entire time interval.

%\begin{figure}
%\begin{center}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_fom_t5_basis2.pdf}
%\caption{FOM} 
%\end{subfigure}
%%\begin{subfigure}[t]{0.95\textwidth}
%%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c2_t10.pdf}
%%\caption{\methodAcronym, $\Delta T = 0.2$} 
%%\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_lspg_t5_basis2.pdf}
%\caption{LSPG} 
%\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c5_t5_basis2.pdf}
%\caption{\methodAcronym, $\Delta T = 0.5$} 
%\end{subfigure}
%%\begin{subfigure}[t]{0.95\textwidth}
%%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c10_t10.pdf}
%%\caption{\methodAcronym, $\Delta T = 2.0$} 
%%\end{subfigure}
%\caption{Field solutions for the vorticity at $t = 5.0$}
%\label{fig:cav_snapshots}
%\end{center}
%\end{figure}

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/fom.png}
\caption{FOM}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/lspg.png}
\caption{LSPG}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window5.png}
\caption{\methodAcronym, $\Delta T = 0.5$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window10.png}
\caption{\methodAcronym, $\Delta T = 1.0$}
\end{subfigure}
\caption{Vorticity snapshots from the FOM and ROM simulations at $t=5.0$.} 
\label{fig:cav_snapshots}
\end{center}
\end{figure}
